---
title: "Bootstrap for regression lab"
author: "Shaoran Sun"
date: "May 3, 2016"
output: pdf_document
---

#The bootstrap for simple linear regression
##1
The estimated slope is the relationship between SAT and college GPA: for every unit of increasement of SAT, there is 0.0018 unit of increasement of college GPA.

##2
When the index is in the form of a decimal point like: "x.y, or 77.2", then the value is repeated more than once.

##3
The 95% bootstrap interval for the slope of the population regression line is \textbf{(0.001076773, 0.002550616)}

#Duality between confidence intervals and hypothesis tests
##4
Less than 0.05, because 0 is not within the interval.

So we can reject the null hypothesis, and conlcude that $\beta_1 \neq 0$

##5
The 99% bootstrap interval for the slope of the population regression line is \textbf{(0.0008904364, 0.002740267)}

Less than 0.01, because 0 is not within the interval.

So we can reject the null hypothesis, and conlcude that $\beta_1 \neq 0$

##6
When theta is one of the bounds of the interval.

#The bootstrap for multiple regression
##7
CollGPA = 0.00123\*SAT+0.376\*HSGPA+0.0227\*Rec

##8
For SAT: (0.0005905682, 0.001961265)

For HSGPA: (0.1399587, 0.6003998)

For Rec: (-0.09318261, 0.1284393)

##9
$p$-values for both thetas of SAT and HSGPA are less than 0.05, because 0 is not within the interval. So we can reject the null hypothesis, and conlcude that $$\beta_1 \neq 0$$ $$\beta_2 \neq 0$$

$p$-value for theta of Rec, however, is greater than 0.05, because 0 is within the interval. So we can fail to reject the null hypothesis, and conlcude that $$\beta_3 = 0$$

##10
For $\beta_1 \& \beta_2$, no matter how wide/loose the bootstrap interval is, 0 is never within the interval. 

For $\beta_3$, no matter how narrow/strict the bootstrap interval is, 0 is always within the interval. 

#Lab Summary
We have the following result initially:

CollGPA = 0.00123\*SAT+0.376\*HSGPA+0.0227\*Rec

After the bootstrap interval test: 

$p$-values for both thetas of SAT and HSGPA are less than 0.05. So we can conlcude that $\beta_1 \neq 0 \& \beta_2 \neq 0$

$p$-value for theta of Rec, however, is greater than 0.05. So we can conlcude that $\beta_3 = 0$

Hence, we can conlcude that variable $Rec$ is not significant, and SAT and HSGPA are significant. We have very strong evidence that SAT is significant, and relatively strong evidence that HSGPA is significant.

\newpage

#Appendix
##1
```{r}
### create our data and calculate thetahat, the slope of the regression line
GPA <- read.delim("~/Desktop/3480 Lab/GPA.txt")
oursample = GPA
regr=lm(CollGPA ~ SAT, data=oursample)
summary(regr)
thetahat = lm(CollGPA ~ SAT, data=oursample)$coeff[2]
thetahat
thetahat.b = rep(NA,1000)
for (i in 1:1000) {
### draw the bootstrap sample and calculate thetahat.b
index = 1:100
bootindex = sample(index, 100, replace=T)
bootsample = oursample[bootindex,]
thetahat.b[i] = lm(CollGPA ~ SAT, data=bootsample)$coeff[2]
}
hist(thetahat.b)
quantile(thetahat.b, .025); quantile(thetahat.b, .975)
```

##2
```{r}
### draw the bootstrap sample
index = 1:100
bootindex = sample(index, 100, replace=T)
bootsample = oursample[bootindex,]
bootsample
```

##5
```{r}
quantile(thetahat.b, .005); quantile(thetahat.b, .995)
```

##7
```{r}
GPAfull <- read.delim("~/Desktop/3480 Lab/GPAfull.txt")
full.reg=lm(CollGPA ~ SAT + HSGPA + Rec, data=GPAfull)
summary(full.reg)
```

##8
```{r}
### create our data
oursample = GPAfull
SATthetahat = lm(CollGPA ~ SAT + HSGPA + Rec, data=oursample)$coeff[2]
HSGPAthetahat = lm(CollGPA ~ SAT + HSGPA + Rec, data=oursample)$coeff[3]
Recthetahat = lm(CollGPA ~ SAT + HSGPA + Rec, data=oursample)$coeff[4]
SATthetahat; HSGPAthetahat; Recthetahat
SATthetahat.b = rep(NA,1000); HSGPAthetahat.b = rep(NA,1000); Recthetahat.b = rep(NA,1000)
for (i in 1:1000) {
### draw the bootstrap sample and calculate thetahat.b
index = 1:100
bootindex = sample(index, 100, replace=T)
bootsample = oursample[bootindex,]
SATthetahat.b[i] = lm(CollGPA ~ SAT + HSGPA + Rec, data=bootsample)$coeff[2]
HSGPAthetahat.b[i] = lm(CollGPA ~ SAT + HSGPA + Rec, data=bootsample)$coeff[3]
Recthetahat.b[i] = lm(CollGPA ~ SAT + HSGPA + Rec, data=bootsample)$coeff[4]
}
par(mfrow=c(1,3))
hist(SATthetahat.b); hist(HSGPAthetahat.b); hist(Recthetahat.b)
quantile(SATthetahat.b, .025); quantile(SATthetahat.b, .975)
quantile(HSGPAthetahat.b, .025); quantile(HSGPAthetahat.b, .975)
quantile(Recthetahat.b, .025); quantile(Recthetahat.b, .975)
```

##10
```{r}
quantile(SATthetahat.b, .005); quantile(SATthetahat.b, .995)
quantile(HSGPAthetahat.b, .005); quantile(HSGPAthetahat.b, .995)
quantile(Recthetahat.b, .05); quantile(Recthetahat.b, .95)
```