---
title: "STAT 3480 Lab8"
author: "Shaoran Sun"
date: "April 26, 2016"
output: pdf_document
---

# Part 1
$H_0$: There is no association between the row and the column factors. (Gender and sport preference are independent)

$H_a$: There is association between the row and the column factors. (Gender and sport preference are dependent)

# Part 2
No. Because the expected counts of every cell in the table are less than 5.

\begin{center}

\begin{tabular}{c|c|c|c|c}
\hline
 & basketball & football & other & total\\
\hline
boys  & 1 & 2 & 0 & 3    \\
girls & 1 & 0 & 1 & 2  \\
\hline
total & 2 & 2 & 1 & 5 \\
\hline
\end{tabular}


Table of Expected Counts

\begin{tabular}{c|c|c|c|c}
\hline
 & basketball & football & other & total\\
\hline
boys  & 6/5 & 6/5 & 3/5 & 3    \\
girls & 4/5 & 4/5 & 2/5 & 2  \\
\hline
total &  2  &  2  &  1  & 5 \\
\hline
\end{tabular}
\end{center}

# Part 3
We need at least 63 students. 

Since the least expected count in the table is "girls other", we need the expected count of this cell greater than 5, so number others should be greater than $5\div\frac{2}{5}$, which is 12.5. This is $1/5$ of the total students. Hence, we have a total number of students of $12.5/(1/5)=62.5$

# Part 4
$\chi^2_{obs}$ = 2.9167

# Part 5
There are 10 rows in the table.

\begin{center}
\begin{tabular}{l*{2}|{c}r}
\hline
boys  &   girls  \\
\hline
B1 B2 F1&F2 O1 \\
B1 B2 F2&F1 O1 \\
B1 B2 O1&F1 F2 \\
B1 F1 F2&B2 O1 \\
B1 F1 O1&B2 F2 \\
B1 F2 O1&B2 F1 \\
B2 F1 F2&B1 O1 \\
B2 F1 O1&B1 F2 \\
B2 F2 O1&B1 F1 \\
F1 F2 O1&B1 B2 \\
\end{tabular}
\end{center}

# Part 6
\begin{center}
\begin{tabular}{l*{2}|{c}r}
\hline
boys  &   girls  \\
B F O & B F O \\
\hline
2 1 0& 0 1 1\\
2 1 0& 0 1 1\\
2 0 1& 0 2 0\\
1 2 0& 1 0 1\\
1 1 1& 1 1 0\\
1 1 1& 1 1 0\\
1 2 0& 1 0 1\\
1 1 1& 1 1 0\\
1 1 1& 1 1 0\\
0 2 1& 2 0 0\\
\end{tabular}
\end{center}

# Part 7
\begin{center}
\begin{tabular}{l*{3}|{c}r}
\hline
boys  &   girls  & test statistic $\chi^2$\\
B F O & B F O \\
\hline
2 1 0& 0 1 1&2.9167\\
2 1 0& 0 1 1&2.9167\\
2 0 1& 0 2 0&5\\
1 2 0& 1 0 1&2.9167\\
1 1 1& 1 1 0&0.83333\\
1 1 1& 1 1 0&0.83333\\
1 2 0& 1 0 1&2.9167\\
1 1 1& 1 1 0&0.83333\\
1 1 1& 1 1 0&0.83333\\
0 2 1& 2 0 0&5\\
\end{tabular}
\end{center}


# Part 8
$\chi^2_{obs}$ is 2.9167. Out of the ten permutations, 6 are greater or equal to 2.9167. Hence, the $p$-value is 6/10 = 0.6, which is greater than 0.05. Therefore we fail to reject the null hypothesis and conclude that there is not enough evidence to show gender and sport preference are dependent.

# Part 9
$\chi^2_{obs}$ is 8.744026. Out of the 1000 random permutations, the $p$-value is 0.013, which is less than 0.05. Therefore we reject the null hypothesis and conclude that gender and sport preference are dependent.


# Part 10
Yes. The least expected count is 10.11, which is greater than 5. So we can apply traditional method. We have the same test statistic of 8.744, and similar $p$-value of 0.013, which is still less than 0.05. Therefore we still reject the null hypothesis and conclude that gender and sport preference are dependent.




\begin{center}

\begin{tabular}{c|c|c|c|c}
\hline
 & basketball & football & other & total\\
\hline
boys  & 22 & 40 & 12 & 74    \\
girls & 30 & 17 & 11 & 58  \\
\hline
total & 52 & 57 & 23 & 132 \\
\hline
\end{tabular}


Table of Expected Counts

\begin{tabular}{c|c|c|c|c}
\hline
 & swimming & soccer & track & total\\
\hline
boys  &  4&4  &4  &    12 \\
girls &4  &4  &3  &   11\\
\hline
total &   8 &  8  &  7  & 23 (grand total = 132) \\
\hline
\end{tabular}
\end{center}


#Appendix
##4
```{r}

A = matrix( c(1,2,0,1,0,1), # the data elements 
            nrow=2,              # number of rows 
            ncol=3,              # number of columns 
            byrow = TRUE)        # fill matrix by rows 

dimnames(A) = list( c("boys", "girls"),         # row names 
                    c("basketball", "football", "other")) # column names 
chisq.test(A)
```

##7
```{r}
chisq.test(rbind(c(2 ,1 ,0),c( 0, 1, 1)))
chisq.test(rbind(c(2, 1, 0),c( 0, 1, 1)))
chisq.test(rbind(c(2 ,0 ,1),c( 0 ,2 ,0)))
chisq.test(rbind(c(1, 2 ,0),c( 1 ,0 ,1)))
chisq.test(rbind(c(1, 1 ,1),c(  1 ,1 ,0)))
chisq.test(rbind(c(1, 1 ,1),c(  1 ,1 ,0)))
chisq.test(rbind(c(1, 2 ,0),c(  1 ,0 ,1)))
chisq.test(rbind(c(1, 1 ,1),c(  1 ,1 ,0)))
chisq.test(rbind(c(1, 1 ,1),c(  1 ,1 ,0)))
chisq.test(rbind(c(0, 2 ,1),c(  2 ,0, 0)))
```

##9
```{r}
### Make observed contingency table and calculate stat
Row1 = c(22,40,12); Row2 = c(30,17,11)
Table = rbind(Row1,Row2)
teststat.obs = chisq.test(Table)$statistic
teststat.obs

### create the prefernce data and the gender data
preference = c( rep("B",52), rep("F",57), rep("O",23))
gender = c( rep("boy",74), rep("girl",58) )
table(preference); table(gender)
y = preference; x = gender
teststat = rep(NA, 1000)
for(i in 1:1000) {
### randomly "shuffle" the y data between the x groups
ySHUFFLE = sample(y)
### compute chi-square stat for the shuffled data
TableSHUFFLE = table(x,ySHUFFLE)
teststat[i] = chisq.test(TableSHUFFLE)$statistic
}
### calculate the approximate p-value
sum(teststat >= teststat.obs)/1000
sum(teststat >= teststat.obs)
```

##10
```{r}
Row1 = c(22,40,12); Row2 = c(30,17,11)
Table = rbind(Row1,Row2)
chisq.test(Table)
```